On existence of integral point sets and their diameter bounds \footnote{ This work was carried out at Voronezh State University and supported by the Russian Science Foundation grant 19-11-00197. }

### 2019-11-05 pps-linear-bound/pps-linear-lower-bound.tex
[J220] Report 1, remark 2
I preferred the first suggested option, i.e. replacing c_j -> 2c_J
[J220] Report 2, starting the major revision
The referree has advices to try to better guide the reader through the long list of technical auxiliary results.
So, all the auxiliary results concerning the intersection of an integral point set with a line segment
have been separated into a new section.

@@

Now we need to estimate the cardinality of an intersection of an integral point set with a line segment.
Assuming that the planar integral point sets contains many collinear points,
the following result holds.
\begin{theorem}~\cite[Theorem 4]{kurz2008minimum}
	For $\delta > 0$, $\varepsilon > 0$, and $P\in\mathfrak{M}(2,n)$ with
	at least $n^\delta$ collinear points there exists a $n_0 (\varepsilon)$
	such that for all $n \geq n_0 (\varepsilon)$ we have
	\begin{equation}
		\operatorname{diam} P \geq n^{\frac{\delta}{4 \log 2(1+\varepsilon)}\log \log n}
		.
	\end{equation}
\end{theorem}
However, the estimate is rather unsuitable for our needs,
as it does not provide the values of all the constants.
To obtain the needed estimate, we now prove a generalization of~\cite[lemma 3]{our-vmmsh-2018}.

\begin{lemma}
	Let $\Delta$ be a straight line segment, $|\Delta| = b$ and $M \in \mathfrak{M}(2,\mathbb{N})$.
	Let $\#(\Delta \cap M) = t$.
	Then
	\begin{equation}
		\label{eq:estimate_for_segment_length}
		b\geq \frac{2}{3}t^{3/2}-\frac{3}{2}t+\frac{5}{6}t^{1/2}
		.
	\end{equation}
\end{lemma}

This lemma leads to the following proposition.

@@

	Next we define the coordinates of the points as following:
	\begin{equation}
		M_{J\pm} =\left(\pm\frac{b_J}{2}, 0\right)
		,
		~~
		N   =\left(0, \frac{\sqrt{a}}{2}\right)
		.
	\end{equation}
	Then the distances are:
	\begin{multline}
		|N M_{J\pm}|
		=
		\left(\frac{b_J^2}{4} + \frac{a}{4}\right)^{1/2}
		=
		\frac{1}{2}\left(\left(\frac{c_J - a/c_J}{2}\right)^2 + a\right)^{1/2}
		=
		\\=
		\frac{1}{2}\left( \left(\frac{c_J}{2}\right)^2 - \frac{a}{2} + \left(\frac{a}{c_j}\right)^2 + a\right)^{1/2}
		=
		\frac{1}{2}\left( \left(\frac{c_J}{2}\right)^2 + \frac{a}{2} + \left(\frac{a}{c_j}\right)^2    \right)^{1/2}
		=
		\\=
		\frac{1}{2}\left(\left(\frac{c_J + a/c_J}{2}\right)^2\right)^{1/2}
		=
		\frac{1}{2}\left(\frac{c_J + a/c_J}{2}\right)
		=
		\frac{g_J}{2}
		\in\mathbb{N}
		,
	\end{multline}
	\begin{equation}
		|M_{J_1 \pm}  M_{J_2 \pm}|
		=
		\left|\frac{b_{J_1}}{2} \pm \frac{b_{J_2}}{2}\right|
		=
		\left|\frac{b_{J_1} \pm b_{J_2}}{2}\right|
		\in\mathbb{N}
		.
	\end{equation}
	In particular, for $H = \{k-1\}$ we obtain $C_H = 2^{2^{k-1}}+1$ and
	\begin{equation}
		b_H =
		\left( 2^{2^{k-1}}+1 - \frac{a}{2^{2^{k-1}}+1} \right)/2
		=
		\left(2^{2^{k-1}}+1 - \left(  2^{2^{k-1}}-1 \right) \right)/2
		=
		1
		.
	\end{equation}
	Thus, one of the distances is
	\begin{equation}
		|M_{H+}  M_{H-}|
		=
		\left|\frac{b_{H}}{2} - \frac{-b_{H}}{2}\right|
		=
		\left|\frac{1}{2} - \frac{-1}{2}\right|
		= 1
		.
	\end{equation}

>>>>

\section{Collinear subsets of integral point sets}
Integral point sets with many collinear ponts are very dominating examples.
In this section we estimate the cardinality of an intersection of an integral point set with a line segment.

Assuming that the planar integral point set contains many collinear points,
the following result holds.
\begin{theorem}~\cite[Theorem 4]{kurz2008minimum}
	For $\delta > 0$, $\varepsilon > 0$, and $P\in\mathfrak{M}(2,n)$ with
	at least $n^\delta$ collinear points there exists a $n_0 (\varepsilon)$
	such that for all $n \geq n_0 (\varepsilon)$ we have
	\begin{equation}
		\operatorname{diam} P \geq n^{\frac{\delta}{4 \log 2(1+\varepsilon)}\log \log n}
		.
	\end{equation}
\end{theorem}
However, the estimate is rather unsuitable for our needs,
as it does not provide the values of all the constants.
To obtain the needed estimate, we now prove a generalization of~\cite[lemma 3]{our-vmmsh-2018}.

\begin{lemma}
	\label{lem:line_segment_cardinality}
	Let $\Delta$ be a straight line segment, $|\Delta| = b$ and $M \in \mathfrak{M}(2,\mathbb{N})$.
	Let $\#(\Delta \cap M) = t$.
	Then
	\begin{equation}
		\label{eq:estimate_for_segment_length}
		b\geq \frac{2}{3}t^{3/2}-\frac{3}{2}t+\frac{5}{6}t^{1/2}
		.
	\end{equation}
\end{lemma}

In~\cite{kurz2008bounds}, thorough investigation of planar integral point sets with diameter at most 10 000
is carried out;
thus, we can restrict our attention to intersections of planar integral point sets with lne segments of length more than 10 000.

Lemma~\ref{lem:line_segment_cardinality} leads to the following proposition.

Concluding the section, we remark that
Corollary~\ref{cor:only_one_distance_1_on_straight_line}
and
Proposition~\ref{obs:estimate_points_on_straight_line}
are the only results of this section
that are used in the proofs below;
all the other are auxiliary.

@@

Now we need to estimate the cardinality of an intersection of an integral point set with a line segment.

@@

	Next we define the coordinates of the points as following:
	\begin{equation}
		M_{J\pm} =\left(\pm\frac{b_J}{2}, 0\right)
		,
		~~
		N   =\left(0, \frac{\sqrt{a}}{2}\right)
		.
	\end{equation}
	Then the distances are:
	\begin{multline}
		|N M_{J\pm}|
		=
		\left(\frac{b_J^2}{4} + \frac{a}{4}\right)^{1/2}
		=
		\frac{1}{2}\left(\left(\frac{c_J - a/c_J}{2}\right)^2 + a\right)^{1/2}
		=
		\\=
		\frac{1}{2}\left( \left(\frac{c_J}{2}\right)^2 - \frac{a}{2} + \left(\frac{a}{2c_J}\right)^2 + a\right)^{1/2}
		=
		\frac{1}{2}\left( \left(\frac{c_J}{2}\right)^2 + \frac{a}{2} + \left(\frac{a}{2c_J}\right)^2    \right)^{1/2}
		=
		\\=
		\frac{1}{2}\left(\left(\frac{c_J + a/c_J}{2}\right)^2\right)^{1/2}
		=
		\frac{1}{2}\left(\frac{c_J + a/c_J}{2}\right)
		=
		\frac{g_J}{2}
		\in\mathbb{N}
		,
	\end{multline}
	\begin{equation}
		|M_{J_1 \pm}  M_{J_2 \pm}|
		=
		\left|\frac{b_{J_1}}{2} \pm \frac{b_{J_2}}{2}\right|
		=
		\left|\frac{b_{J_1} \pm b_{J_2}}{2}\right|
		\in\mathbb{N}
		.
	\end{equation}
	In particular, for $H = \{k-1\}$ we obtain $C_H = 2^{2^{k-1}}+1$ and
	\begin{equation}
		b_H =
		\left( 2^{2^{k-1}}+1 - \frac{a}{2^{2^{k-1}}+1} \right)/2
		=
		\left(2^{2^{k-1}}+1 - \left(  2^{2^{k-1}}-1 \right) \right)/2
		=
		1
		.
	\end{equation}
	Thus, one of the distances is
	\begin{equation}
		|M_{H+}  M_{H-}|
		=
		\left|\frac{b_{H}}{2} - \frac{-b_{H}}{2}\right|
		=
		\left|\frac{1}{2} - \frac{-1}{2}\right|
		= 1
		.
	\end{equation}
